---
title: "gvc_agora_opentargets"
format:
  html:
    df-print: paged
    toc: true
execute: 
  warning: false
---

## Setup environment

```{r}
#| output: false
library(tidyverse)
library(janitor)
library(broom)
library(readxl)
library(jsonlite)

library(gprofiler2)

theme_set(theme_bw())

set.seed(666)
```

## Read and prep data

### GVC

Genes within 1Mb window of (each side of?) GVC loci from Fanny:

```{r}
gvc <- read_xlsx("GVC_1Mb_comparison_050224.xlsx") |>
  clean_names() |> 
  separate(gene_id, c("gene_id", "version")) |>
  select(-version, -agora_nominated_list, -opentarget_info)

gvc
```

```{r}
gvc.genes <- gvc |> distinct(gene_id, .keep_all = TRUE) |> select(gene_id, gene_symbol) |> arrange(gene_symbol)

gvc.genes
```

### Agora

Alzheimer's disease gene prioritization scores from [Agora](https://www.synapse.org/Synapse:syn25741025):

```{r}
ago1 <- read_json("syn25741025.overall_scores.json", simplifyVector = TRUE) |> as_tibble()

ago1
```

Alzheimer's disease genes (AMPAD Agora) from Fanny:

```{r}
ago2 <- read_csv("AMPAD_agora_032124_gene-list.csv")
ago2
```

```{r}
ago <- ago1 |> filter(hgnc_symbol %in% ago2$`Gene Symbol`)
```

### OpenTargets

Alzheimer's disease gene prioritization scores from OpenTargets:

```{r}
ot <- read_tsv("OT-MONDO_0004975-associated-targets-6_4_2024-v24_03.tsv", show_col_types = FALSE, na = "No data")

ot
```

Add Ensembl Gene IDs (WTF!):

```{r}
otcols <- colnames(ot)
otensg <- gconvert(
  query = ot$symbol,
  organism = "hsapiens",
  target= "ENSG",
  mthreshold = Inf,
  filter_na = TRUE) |> 
  mutate(input_number = as.character(input_number)) |>
  left_join(ot |> rownames_to_column(var = "input_number"), by = "input_number") |> 
  select(ensembl_gene_id = target, otcols)

otensg
```

## Annotate GVC genes with Agora and OpenTargets scores

```{r}
sum(gvc.genes$gene_id %in% ago$ensembl_gene_id)
```

```{r}
sum(gvc.genes$gene_id %in% otensg$ensembl_gene_id)
```

Arrange by Agora's `genetics_score` and OpenTargets' `otGeneticsPortal`:

```{r}
d1 <- gvc.genes |>
  left_join(ago, by = join_by(gene_id == ensembl_gene_id)) |> 
  left_join(otensg, by = join_by(gene_id == ensembl_gene_id)) |> 
  arrange(desc(genetics_score), desc(otGeneticsPortal))

d1
```

## Overlap between GVC genes and Agora and OpenTargets genes

```{r}
x = list(
  "GVC" = gvc.genes$gene_id,
  "Agora" = ago$ensembl_gene_id,
  "OpenTargets" = otensg$ensembl_gene_id
)
```

```{r}
#| fig-width: 6
#| fig-height: 6
library(VennDiagram)
grid.newpage()
v <- venn.diagram(
  x,
  fill = c("#FF0000", "#00FF00", "#0000FF"),
  filename = NULL)
grid.draw(v)
```

```{r}
p <- get.venn.partitions(x)
p
```

## Perform ORA of overlap genes

### GVC ∩ Agora ∩ OpenTargets

```{r}
query <- p |> 
  filter(..set.. == "GVC∩Agora∩OpenTargets") |>
  unnest(..values..) |>
  select(gene_id = ..values..) |>
  left_join(ago, by = join_by(gene_id == ensembl_gene_id)) |> 
  left_join(otensg, by = join_by(gene_id == ensembl_gene_id)) |> 
  arrange(desc(genetics_score), desc(otGeneticsPortal)) |> 
  distinct(gene_id) |>
  pull(gene_id)
```

```{r}
gostres <- gost(query = query,
                organism = "hsapiens",
                domain_scope = "annotated",
                exclude_iea = TRUE,
                ordered_query = TRUE,
                significant = TRUE,
                user_threshold = 0.005,
                correction_method = "fdr")

gostres$result |> select(term_name, term_id, source, everything())
```

```{r}
gostplot(gostres, capped = FALSE, interactive = TRUE)
```

```{r}
# save overlap gene ids for later
overlap_gene_ids <- query
```

### (Agora ∩ OpenTargets) ∖ (GVC)

```{r}
query <- p |>
  filter(..set.. == "(Agora∩OpenTargets)∖(GVC)") |> 
  unnest(..values..) |>
  select(gene_id = ..values..) |>
  left_join(ago, by = join_by(gene_id == ensembl_gene_id)) |> 
  left_join(otensg, by = join_by(gene_id == ensembl_gene_id)) |> 
  arrange(desc(genetics_score), desc(otGeneticsPortal)) |> 
  distinct(gene_id) |>
  pull(gene_id)
```

```{r}
gostres <- gost(query = query,
                organism = "hsapiens",
                domain_scope = "annotated",
                exclude_iea = TRUE,
                ordered_query = TRUE,
                significant = TRUE,
                user_threshold = 0.005,
                correction_method = "fdr")

gostres$result |> select(term_name, term_id, source, everything())
```

```{r}
gostplot(gostres, capped = FALSE, interactive = TRUE)
```

### (GVC ∩ OpenTargets) ∖ (Agora)

```{r}
query <- p |>
  filter(..set.. == "(GVC∩OpenTargets)∖(Agora)") |> 
  unnest(..values..) |>
  select(gene_id = ..values..) |>
  left_join(ago, by = join_by(gene_id == ensembl_gene_id)) |> 
  left_join(otensg, by = join_by(gene_id == ensembl_gene_id)) |> 
  arrange(desc(genetics_score), desc(otGeneticsPortal)) |> 
  distinct(gene_id) |>
  pull(gene_id)
```

```{r}
gostres <- gost(query = query,
                organism = "hsapiens",
                domain_scope = "annotated",
                exclude_iea = TRUE,
                ordered_query = TRUE,
                significant = TRUE,
                user_threshold = 0.005,
                correction_method = "fdr")

gostres$result |> select(term_name, term_id, source, everything())
```

```{r}
gostplot(gostres, capped = FALSE, interactive = TRUE)
```

### (OpenTargets) ∖ (GVC ∪ Agora)

```{r}
query <- p |>
  filter(..set.. == "(OpenTargets)∖(GVC∪Agora)") |> 
  unnest(..values..) |>
  select(gene_id = ..values..) |>
  left_join(ago, by = join_by(gene_id == ensembl_gene_id)) |> 
  left_join(otensg, by = join_by(gene_id == ensembl_gene_id)) |> 
  arrange(desc(genetics_score), desc(otGeneticsPortal)) |> 
  distinct(gene_id) |>
  pull(gene_id)
```

```{r}
gostres <- gost(query = query,
                organism = "hsapiens",
                domain_scope = "annotated",
                exclude_iea = TRUE,
                ordered_query = TRUE,
                significant = TRUE,
                user_threshold = 0.005,
                correction_method = "fdr")

gostres$result |> select(term_name, term_id, source, everything())
```

```{r}
gostplot(gostres, capped = FALSE, interactive = TRUE)
```

### (GVC ∩ Agora) ∖ (OpenTargets)

```{r}
query <- p |>
  filter(..set.. == "(GVC∩Agora)∖(OpenTargets)") |> 
  unnest(..values..) |>
  select(gene_id = ..values..) |>
  left_join(ago, by = join_by(gene_id == ensembl_gene_id)) |> 
  left_join(otensg, by = join_by(gene_id == ensembl_gene_id)) |> 
  arrange(desc(genetics_score), desc(otGeneticsPortal)) |> 
  distinct(gene_id) |>
  pull(gene_id)
```

```{r}
gostres <- gost(query = query,
                organism = "hsapiens",
                domain_scope = "annotated",
                exclude_iea = TRUE,
                ordered_query = TRUE,
                significant = TRUE,
                user_threshold = 0.005,
                correction_method = "fdr")

gostres$result |> select(term_name, term_id, source, everything())
```

```{r}
gostplot(gostres, capped = FALSE, interactive = TRUE)
```

### (Agora) ∖ (GVC ∪ OpenTargets)

```{r}
query <- p |>
  filter(..set.. == "(Agora)∖(GVC∪OpenTargets)") |> 
  unnest(..values..) |>
  select(gene_id = ..values..) |>
  left_join(ago, by = join_by(gene_id == ensembl_gene_id)) |> 
  left_join(otensg, by = join_by(gene_id == ensembl_gene_id)) |> 
  arrange(desc(genetics_score), desc(otGeneticsPortal)) |> 
  distinct(gene_id) |>
  pull(gene_id)
```

```{r}
gostres <- gost(query = query,
                organism = "hsapiens",
                domain_scope = "annotated",
                exclude_iea = TRUE,
                ordered_query = TRUE,
                significant = TRUE,
                user_threshold = 0.005,
                correction_method = "fdr")

gostres$result |> select(term_name, term_id, source, everything())
```

```{r}
gostplot(gostres, capped = FALSE, interactive = TRUE)
```

### (GVC) ∖ (Agora ∪ OpenTargets)

```{r}
query <- p |>
  filter(..set.. == "(GVC)∖(Agora∪OpenTargets)") |> 
  unnest(..values..) |>
  select(gene_id = ..values..) |>
  left_join(ago, by = join_by(gene_id == ensembl_gene_id)) |> 
  left_join(otensg, by = join_by(gene_id == ensembl_gene_id)) |> 
  arrange(desc(genetics_score), desc(otGeneticsPortal)) |> 
  distinct(gene_id) |>
  pull(gene_id)
```

```{r}
gostres <- gost(query = query,
                organism = "hsapiens",
                domain_scope = "annotated",
                exclude_iea = TRUE,
                ordered_query = TRUE,
                significant = TRUE,
                user_threshold = 0.005,
                correction_method = "fdr")

gostres$result |> select(term_name, term_id, source, everything())
```

```{r}
gostplot(gostres, capped = FALSE, interactive = TRUE)
```

## Perform ORA of GVC genes sorted by Agora or OpenTargets genetics scores

### Agora

```{r}
query <- d1 |> arrange(desc(genetics_score)) |> distinct(gene_id) |> pull(gene_id)

gostres <- gost(query = query,
                organism = "hsapiens",
                domain_scope = "annotated",
                exclude_iea = TRUE,
                ordered_query = TRUE,
                significant = TRUE,
                user_threshold = 0.005,
                correction_method = "fdr")

gostres$result |> select(term_name, term_id, source, everything())
```

```{r}
gostplot(gostres, capped = FALSE, interactive = TRUE)
```

### OpenTargets

```{r}
query <- d1 |> arrange(desc(otGeneticsPortal)) |> distinct(gene_id) |> pull(gene_id)

gostres <- gost(query = query,
                organism = "hsapiens",
                domain_scope = "annotated",
                exclude_iea = TRUE,
                ordered_query = TRUE,
                significant = TRUE,
                user_threshold = 0.005,
                correction_method = "fdr")

gostres$result |> select(term_name, term_id, source, everything())
```

```{r}
gostplot(gostres, capped = FALSE, interactive = TRUE)
```

## Correlation of Agora and OpenTargets scores in GVC genes

```{r}
d1 |> nrow()

d1 |> drop_na(genetics_score, otGeneticsPortal) |> nrow()

d1 |> drop_na(genetics_score, otGeneticsPortal) |>
  summarize(cor = tidy(cor.test(genetics_score, otGeneticsPortal, method="kendall"))) |>
  unnest(cor)
```

```{r}
d1 |> nrow()

d1 |> drop_na(target_risk_score, globalScore) |> nrow()

d1 |> drop_na(target_risk_score, globalScore) |>
  summarize(cor = tidy(cor.test(target_risk_score, globalScore, method="kendall"))) |>
  unnest(cor)
```

## Correlation of Agora and OpenTargets scores overall

```{r}
d2 <- ago |> 
  left_join(otensg, by = "ensembl_gene_id") |> 
  arrange(desc(genetics_score), desc(otGeneticsPortal))

d2
```

```{r}
d2 |> nrow()

d2 |> drop_na(genetics_score, otGeneticsPortal) |> nrow()

d2 |> drop_na(genetics_score, otGeneticsPortal) |>
  summarize(cor = tidy(cor.test(genetics_score, otGeneticsPortal, method="kendall"))) |>
  unnest(cor)
```

```{r}
d2 |> nrow()

d2 |> drop_na(target_risk_score, globalScore) |> nrow()

d2 |> drop_na(target_risk_score, globalScore) |>
  summarize(cor = tidy(cor.test(target_risk_score, globalScore, method="kendall"))) |>
  unnest(cor)
```

## Print environment

```{r}
sessioninfo::session_info()
```
